---
title: "Ontology Manipulation"
output: html_notebook
---

For classifying cells and for comparing a hierarchically clustered tree to the 'ground truth' cell ontology tree.

```{r}
library(ontologyIndex)
library(ontologyPlot)
library(ontologySimilarity)
library(tidyverse)
library(here)

co = get_ontology('https://raw.githubusercontent.com/obophenotype/cell-ontology/master/cl-basic.obo', extract_tags='everything')
```

```{r}
tm.facs.metadata = read_csv(here("data", "TM_facs_metadata.csv"))
tm.droplet.metadata = read_csv(here("data", "TM_droplet_metadata.csv"))
```

```{r}
tm.ids = unique(c(tm.facs.metadata %>% pull(cell_ontology_id), tm.droplet.metadata %>% pull(cell_ontology_id)))
```

What are the immune cell types in Tabula Muris?

```{r}
co$name[grep(pattern = 'leukocyte', x = co$name)]
onto_plot(co, terms=intersection_with_descendants(co,"CL:0000738",tm.ids))
```

How do those fit into the bigger hieirarchy?

```{r}
onto_plot(co, terms=intersection_with_descendants(co,"CL:0000738",get_ancestors(co,tm.ids)))
```

How do these fit into the full lineage? While there are other, high-order terms, it's most natural for the mouse to begin with Animal Cell.

```{r, fig.width = 20}
#onto_plot(co, terms=intersection_with_descendants(co,"CL:0000548",get_ancestors(co, c("CL:0000236", "CL:0002420"))))
pdf('~/Desktop/all_ontology.pdf')
onto_plot(co, terms=intersection_with_descendants(co,"CL:0000548",get_ancestors(co, tm.ids)))
dev.off()
```

```{r}
co$name[get_descendants(co, c("CL:0000542"))]
```

A similarity matrix between ontology terms is built into ontologySimilarity. I'm not sure exactly what it is computing.

```{r}
sim_mat <- get_sim_grid(ontology=co, term_sets=list(intersection_with_descendants(co,"CL:0000738",tm.ids)))
sim_mat
```

Since an ontology is just a collection of `is_a` relationships, the most straightforward distance between nodes will be a function of the two sets of attributes posessed by the nodes.

The number of unshared attributes (the set disjunction) can be easily computed as (union - intersection) of the ancestor sets. In the case of a tree, this is just the tree distance.

```{r}
ont_dist <- function(ontology, id1, id2){
  anc1 = get_ancestors(ontology, id1)
  anc2 = get_ancestors(ontology, id2)
  length(union(anc1, anc2)) - length(intersect(anc1, anc2))
}
```

Since it's not a tree, though, the distance is somewhat larger.

```{r}
plot_common_tree <- function(t1, t2){
  onto_plot(co, terms = intersection_with_descendants(co, "CL:0000548", union(get_ancestors(co, t1), get_ancestors(co, t2))))
}
plot_common_tree("CL:0000818", "CL:0000914")
ont_dist(co, "CL:0000818", "CL:0000914")
```

Given a distance function between nodes, we can assign a distance between trees by asking for the distortion of the pairwise-distance matrices. This is equivalent to embedding the space of trees into the metric space of distance matrices, and taking the $L^1$ distance there.

Popular methods for comparing trees (or DAGs) also include some kind of mutation distance, based on the number of elementary mutations (such as the contraction of an edge) required to turn one tree into another.

The Robinson-Foulds distance is given by considering the bipartitions introduced by cutting a tree, and comparing those sets of bipartitions. This can also be viewed as an inherited embedding distance, where a tree $T$ is mapped into $2^{2^S}$, the set of subsets of subsets of $S$. Namely, $\gamma(T)$ contains a subset $s \subset S$ if deleting some edge $e$ from $T$ separates the labelled vertices into $\{s, s^c\}$. It is also a mutation distance: construct a graph $\mathcal{T}$ on the space of partially labelled trees, labelled by elements of $S$, where two trees are connected by an edge if one can be produced from the other by collapsing an edge. The RF distance between trees is the graph distance in $\mathcal{T}$. 

Since removing `is_a` relationships will not disconnect the ontology DAG, it is not straightforward to generalize the RF distance.


Let's compute the distance matrix between all ids appearing in TM according to this metric.

```{r}
nids = length(tm.ids)
D = matrix(nrow = nids, ncol = nids)
for (i in 1:nids){
  for(j in 1:nids){
    D[i,j] = ont_dist(co, tm.ids[i], tm.ids[j])
  }
}
```

Cluster with that distance.

```{r, fig.height = 12, fig.width = 18}
rownames(D) = co$name[tm.ids]
colnames(D) = co$name[tm.ids]
d = as.dist(D)

h <- hclust(d, method = "ward.D2")
plot(h, hang = -4)
```


# Gene Ontology

```{r}
data(go)
go$name[grep('surface', go$name)]
onto_plot(go, terms = get_ancestors(go,"GO:0007475"))
```

