---
title: "R Notebook"
output: html_notebook
---

In this post, we will analyze the differential expression of B cells in different compartments of the body. To do this, we will first grab all the cells that were annotated as immune. Since we used the [Cell Ontology](http://obofoundry.org/ontology/cl.html), this is easy as we can take any cells that were annotated as a "leukocyte" or descendant, as this encompasses all immune cells. To enable such analyses, for each cell's annotation we stored both the `cell_ontology_class`, e.g. "leukocyte" and the `cell_ontology_id`. In particular, B cells, which are [lymphoctes, are a descendant (child) of leukocytes](http://www.ontobee.org/ontology/CL?iri=http://purl.obolibrary.org/obo/CL_0000542). The cell ontology ID for [leukocyte](http://purl.obolibrary.org/obo/CL_0000738) is `CL:0000738`. 

First, we load the [`ontologyIndex`](https://www.rdocumentation.org/packages/ontologyIndex/versions/2.4) package for reading ontologies, and grab the latest version of the Cell Ontology from GitHub. Then, we use `get_descendants` to obtain all children terms of "leukocyte," which includes "leukocyte" itself.

```{r}
library(ontologyIndex)

cell_ontology = get_ontology('https://raw.githubusercontent.com/obophenotype/cell-ontology/master/cl-basic.obo',
                             extract_tags='everything')

leukocyte_id = "CL:0000738"
leukocyte_or_descendant = get_descendants(cell_ontology, roots=leukocyte_id)
print(paste("length(leukocyte_or_descendant)", length(leukocyte_or_descendant)))
print(paste("leukocyte_id %in% leukocyte_or_descendant", leukocyte_id %in% leukocyte_or_descendant))
```

Here are examples of some of the children terms of "leukocyte:"

```{r}
head(cell_ontology$name[leukocyte_or_descendant])
```


Now that we have 

