---
title: "B Cells FACS"
output: html_notebook
---
```{r}
library(Seurat)
library(tidyverse)
library(here)
source(here('vignettes', 'bcells', 'common.R'))

load(here('vignettes', 'bcells', 'facs_cd19.Robj'))
```


How many cells do we have per tissue?

```{r}
table(tiss_facs_cd19@meta.data$tissue)
```

```{r}
tissues.enough.cells = table(tiss_facs_cd19@meta.data$tissue) > 30
tissues.enough.cells = tissues.enough.cells[tissues.enough.cells]
tiss_facs_tissues_enough_cells = SubsetData(tiss_facs_cd19, subset.name='tissue', 
                                            accept.value = names(tissues.enough.cells))
sort(unique(tiss_facs_tissues_enough_cells@meta.data$tissue))
table(tiss_facs_tissues_enough_cells@meta.data$tissue)
dim(tiss_facs_tissues_enough_cells@scale.data)
dim(tiss_facs_tissues_enough_cells@scale.data)
```

## Set "Tissue" as the "ident" for further analysis

```{r}
annotation = 'tissue'

annotations = tiss_facs_tissues_enough_cells@meta.data[,annotation]
unique_annotations = unique(annotations)
enumerated_annotations = 0:(length(unique_annotations)-1)

annotation_ident = as.factor(plyr::mapvalues(x = annotations, from = unique_annotations, to = enumerated_annotations))
names(annotation_ident) = names(tiss_facs_tissues_enough_cells@ident)
tiss_facs_tissues_enough_cells@ident = annotation_ident
```


## Hierarchical relationship between tissues based on average expression
```{r}
library(ape)

tiss_facs_tissues_enough_cells = BuildClusterTree(tiss_facs_tissues_enough_cells, do.plot = FALSE)

data.tree = tiss_facs_tissues_enough_cells@cluster.tree[[1]]
data.tree$tip.label = unique_annotations

plot.phylo(x = data.tree, direction = "downwards")
nodelabels()
```

## Differential expression of peripheral tissues vs rest

```{r}
node7_markers = FindMarkersNode(object=tiss_facs_tissues_enough_cells, node=7)
head(node7_markers)
```

```{r}
head(node7_markers)
```
```{r}
write_markers(node7_markers, here('vignettes', 'bcells', 'facs_tissue_node7_markers.csv'))
```


```{r}
top_node7_markers = head(node7_markers, n=12)
node7_genes = rownames(top_node7_markers)
top_node7_markers
```

```{r, fig.width=12, fig.height=8}
RidgePlot(tiss_facs_tissues_enough_cells, node7_genes, group.by='tissue')
```

```{r, fig.width=8}
DotPlot(tiss_facs_tissues_enough_cells, rev(node7_genes), group.by='tissue')
```
## Get subtissues of fat


```{r}
subtissue_if_fat = ifelse(tiss_facs_tissues_enough_cells@meta.data$tissue == "Fat",
                          paste(":", tiss_facs_tissues_enough_cells@meta.data$subtissue), "")
tiss_facs_tissues_enough_cells@meta.data[, 'tissue_subtissue_fat'] = 
  paste0(tiss_facs_tissues_enough_cells@meta.data$tissue, subtissue_if_fat)

tiss_facs_tissues_enough_cells@meta.data[, 'tissue_subtissue'] = 
  paste0(tiss_facs_tissues_enough_cells@meta.data$tissue, ": ", tiss_facs_tissues_enough_cells@meta.data$subtissue)
```


```{r}
DotPlot(tiss_facs_tissues_enough_cells, rev(node7_genes), group.by='tissue_subtissue_fat')
```





## Set "tissue_subtissue_fat" as the "ident" for further analysis

```{r}
annotation = 'tissue_subtissue_fat'

annotations = tiss_facs_tissues_enough_cells@meta.data[,annotation]
unique_annotations = unique(annotations)
enumerated_annotations = 0:(length(unique_annotations)-1)

annotation_ident = as.factor(plyr::mapvalues(x = annotations, from = unique_annotations, to = enumerated_annotations))
names(annotation_ident) = names(tiss_facs_tissues_enough_cells@ident)
tiss_facs_tissues_enough_cells@ident = annotation_ident
```


```{r}
table(tiss_facs_tissues_enough_cells@meta.data$tissue_subtissue_fat)
```

## Hierarchical relationship between tissues and subtissues based on average expression
```{r, fig.height=8}
library(ape)

tiss_facs_tissues_enough_cells = BuildClusterTree(tiss_facs_tissues_enough_cells, do.plot = FALSE)

data.tree = tiss_facs_tissues_enough_cells@cluster.tree[[1]]
data.tree$tip.label = unique_annotations

plot.phylo(x = data.tree, direction = "downwards")
nodelabels()
```


```{r}
tissue_subtissue_of_interest = "Fat: BAT"
enumerated_annotations[unique_annotations == tissue_subtissue_of_interest]
```


```{r}
fat_subtissues = grep("Fat", unique_annotations, value = TRUE)

fat_subtissue_markers = data.frame()
for (fat_subtissue in fat_subtissues){
  ident = enumerated_annotations[unique_annotations == fat_subtissue]
  subtissue_markers = FindMarkers(tiss_facs_tissues_enough_cells, ident.1 = ident)
  subtissue_markers[, "tissue_subtissue"] = fat_subtissue
  subtissue_markers[, "genes"] = rownames(fat_subtissue)
  fat_subtissue_markers = rbind(fat_subtissue_markers, subtissue_markers)
}
```

```{r}
fat_subtissue_markers[, "genes"] = rownames(fat_subtissue_markers)
fat_subtissue_markers %>% group_by(tissue_subtissue) %>% top_n(-10, p_val_adj)
```
```{r}
write_csv(fat_subtissue_markers, here('vignettes', 'bcells', 'facs_fat_subtissue_markers.csv'))
```

```{r}
top_fat_subtissue_markers = fat_subtissue_markers %>% group_by(tissue_subtissue) %>% top_n(-10, p_val_adj)
top_fat_subtissue_genes = sort(unique(top_fat_subtissue_markers$genes))
top_fat_subtissue_genes
```

```{r}
write_csv(top_fat_subtissue_markers, here('vignettes', 'bcells', 'facs_fat_subtissue_markers_top10per.csv'))
```

```{r}
RidgePlot(tiss_facs_tissues_enough_cells,
          features.plot=top_fat_subtissue_genes, 
          group.by='tissue_subtissue_fat')
```

```{r}
DotPlot(tiss_facs_tissues_enough_cells, genes.plot = rev(top_fat_subtissue_genes),
        group.by='tissue_subtissue_fat')
```


## Get differential expression between tissues
```{r}
organ_markers = FindAllMarkers(object = tiss_facs_tissues_enough_cells, only.pos = TRUE, min.pct = 0.25, thresh.use = 0.2)
write_csv(organ_markers, here('vignettes', 'bcells', 'facs_tissue_markers.csv'))
head(organ_markers)
```


```{r}
ggplot(data=organ_markers, aes(x=p_val_adj)) + geom_histogram() + facet_grid(~tissue)
```



```{r}
top_organ_markers = organ_markers %>% group_by(tissue) %>% top_n(-10, p_val_adj)
write_csv(top_organ_markers, here('vignettes', 'bcells', 'facs_tissue_markers_top10_per_tissue.csv'))
genes.to.plot = sort(unique(top_organ_markers$gene))
top_organ_markers
```



```{r, fig.width=12, fig.height=16}
top_organ_markers = organ_markers %>% group_by(tissue) %>% top_n(-3, p_val_adj)
genes.to.plot = sort(unique(top_organ_markers$gene))
print(length(genes.to.plot))

RidgePlot(tiss_facs_tissues_enough_cells, genes.to.plot, group.by='tissue')
```
```{r, fig.width=12}
# Need to reverse gene order so increases alphabetically left to right
DotPlot(tiss_facs_tissues_enough_cells, rev(genes.to.plot), group.by='tissue')
```


```{r, fig.height=40}
nCol = 4
plots = FeaturePlot(tiss_facs_tissues_enough_cells, features.plot = genes.to.plot, 
                    do.return = TRUE, nCol = nCol, no.axes=TRUE, 
                    # Light grey to 
                    cols.use=c('lightgrey', '#008080'))
for (p in plots){
  gene_name = p$labels$title
  gene_tissues = top_organ_markers %>% filter(gene == p$labels$title) %>% select(tissue)
  gene_tissues = as.character(as.vector(gene_tissues$tissue))
  print(paste(gene_name, paste(gene_tissues)))
  title = paste0(gene_name, ' (', paste(gene_tissues, sep=', '), ')')
  p = p + labs(title=title)
  plots[[gene_name]] = p
}
plots.combined <- plot_grid(plotlist = plots, ncol = nCol)
invisible(x = lapply(X = plots.combined, FUN = print))
ggsave('facs_featureplot_top_organ_markers.pdf', height = 40, width=15)
```
